---
title: "Notebook 08"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

# Key Logs (Individual Keys)

Let's read the word-level keylog data into R. 

```{r}
word <- read_csv("../data/keylog_word.csv")
word
```

We are going to test the hypothesis that all students have the
same average gap between finishing one word and starting the
next on the 2nd task. We can grab the data for this using the
following R code:

```{r}
index <- (word$task == 2 & word$gap_after < 2000)
x <- word$gap_after[index]
block <- word$id[index]
```

Then, all of the derived variables can be computed with the
following:

```{r}
xbar <- tapply(x, block, mean)
xbar_all <- mean(x)
s2 <- tapply(x, block, var)
n <- tapply(x, block, length)
K <- length(unique(block))
N <- sum(n)
```

R has a nice syntax where we can do vectorized operations. So,
if we add/multiple two vectors of the same thing, it will do 
these operations component-wise. If we add/multiply a constant
with a vector, it will allow the constant to every entry. For
example, here is the denominator of the F-statistic:

```{r}
sum((n - 1) * s2) / (N - K)
```

Below, write the R code to create the F-statistic from the formula
you derived on today's worksheet. Save the result as an object
named `fstat`:

```{r, question-01}

```

Use the following code to compute the p-value of the F-statistic.
Is the test significant at a 0.001 level?

```{r}
1 - pf(fstat, df1 = (K - 1), df2 = (N - K))
```

As with the other tests we have used, there are build-in R 
functions to do all of this work for us. Here is the code to
run the analysis of variance:

```{r}
summary(aov(x ~ block))
```

There is a lot of information in the output, some of which we
do not need. You should see, though, that there are the
degrees of freedom, the F-statistic, and the p-value. [While
very close, you'll probably notice some numerical instability
of the p-value computation; it is a little different from the
computation (at least on my machine).]
