---
title: "Notebook 06"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

# Key Logs (Individual Keys)

Let's read the keylog data into R again. 

```{r}
kl <- read_csv("../data/keylog.csv")
kl$key[is.na(kl$key)] <- " "
kl <- filter(kl, !is.na(gap1), !is.na(gap2))
kl
```

We start with the same task we looked at last time, the gap
between keys when one is the space bar and the when the key 
is not the space bar. Basically, do we pause more between
words than we do between letters?

```{r}
x <- kl$gap2[kl$task == 1 & kl$key == " " & kl$gap2 < 1000]
y <- kl$gap2[kl$task == 1 & kl$key != "." & kl$gap2 < 1000]
```

Run a T-test using the function `t.test`. You can set the 
variances to be equal (the confidence level is not important)

```{r, question-01}

```

We've already done this before, but now we should understand
all of the output. Would we reject the null hypothesis that
the means of these two gaps are the same?

# User Differences

The code below grabs the duration of key presses of the space
bar for two different students. Here, student "A" and student
"B":

```{r}
x <- kl$gap2[kl$id == "A" & kl$key == " " & kl$gap2 < 1000]
y <- kl$gap2[kl$id == "B" & kl$key == " " & kl$gap2 < 1000]
```

Starting by copying my code above, create two samples of the 
time that it takes to press the space bar for you and the id of
one of your neighbors (ask them their id):

```{r, question-02}

```

Now, run a T-test for the differences in the time that each of
you have after hitting the space bar. Who takes longer? Is the
result significant at the 0.01 level? What is the p-value?

```{r}
t.test(x, y, var.equal = TRUE)
```

Now, using the same data, compute a test of whether each of your
gaps after hitting the space bar have the same variance.

```{r, question-03}

```

Who has more variability? Is this ratio between the variances
significant?

## Words

Below, we will read in a different version of the dataset that 
has keylog information about individual words.

```{r}
word <- read_csv("../data/keylog_word.csv")
word
```

Below, we create new sets of data that correspond to the gap
in milliseconds between words when one word ends a sentence
and when it does not. 

```{r}
x <- word$gap_after[word$task == 1 & stri_detect(word$word, fixed = ".")  & word$gap_after < 2000]
y <- word$gap_after[word$task == 1 & !stri_detect(word$word, fixed = ".") & word$gap_after < 2000]
```

Run a T-test to see whether there is a difference in the pause 
at the end of a sentence and the gap between words.

```{r}
t.test(x, y, var.equal = TRUE)
```

We can look at the words data more next time.
