---
title: "Notebook 15"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

Let's look at an historical dataset of the heights of children
relative to the heights of their parents. This comes from a 
data published by Francis Galton in the 1890s. All heights are
given in inches. Note that these children are all fully grown,
adults.

```{r}
heights <- read_csv("../data/galton_heights.csv")
heights
```

We can use a linear regression to study the height of children
as a function of their mother's height:

```{r}
model <- lm(height ~ mother, data = heights)
summary(model)
```

Using the new multivariate approach, we can add information 
about the height of their mother and the height of their father.

```{r}
model <- lm(height ~ mother + father, data = heights)
summary(model)
```

We can also add variables that correspond to categories. For 
example, we can add a marker for reported gender to the 
regression as follows:

```{r}
model <- lm(height ~ mother + father + gender, data = heights)
summary(model)
```

Notice that the output adds a column called `genderM`, which is
a variable equal to 0 for children labeled as Female and 
1 for children labeled as Male (the baseline is always set to
the alphabetically first category). The estimate, here about 5.22,
provides the expected extra height in inches of male children
relative to the female children.

As a final measurement, the R-square value provides a measurement
of how much variation in the Ys is explained by the model. It
ranges from zero (no explanation) to one (fully explaining the
variable). It is defined as the ratio of the variance of the 
residuals (Y-Xb) divided by the variance of Y. Go back and see
how much better our final model is compared to the first attemps
at explaining the height of the people in the dataset.


