---
title: "Notebook 07"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

In today's notebook, we will make a new small dataset and
then apply the two two-sample tests that we have learned.
The data we will create (it's a bit silly, but easy to 
generate and gives you a quick sense of how to do some 
random sampling) is based off the "Random Page" feature of
Wikipedia. Following the links below will show you a random
Wikipedia page from the English and German versions of the
site, respectively.

- https://en.wikipedia.org/wiki/Special:Random
- https://de.wikipedia.org/wiki/Special:Random

Working either alone or in pairs, construct a dataset by
going to 15 random English pages and 15 random German pages
and recording the number of full-sized images on each page
(do not include maps or small thumbnails). Then, build
two vectors in R with the results as follows (I have filled
in a set of five that you can add to your 15 to get a larger
dataset):

```{r}
x <- c(0, 1, 0, 2, 2)
y <- c(2, 1, 0, 5, 2)
```

Now, all we need to do are run two different statistical
tests/confidence intervals. Use `var.test` below to compare
the variance of the two samples, using the default
null-hypothesis that there is no difference:

```{r, question-01}
var.test(x, y)
```

And then, run `t.test` to test the difference of the means,
using the default null-hypothesis that there is no difference
(you can choose to use the equal variances flag or not, depending
on the previous test's outcome):

```{r, question-02}
t.test(x, y)
```

Make sure that you fully understand what all of the elements
of the output mean.




