---
title: "Notebook 01"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

## Using R

We will be using R and RMarkdown this semester to both apply the 
statistical models that we learn to data and to run simulations
that support our theoretical results. If you have never used R
before that is completely fine! There's a little bit of a
learning curve, but we are mostly either use only relatively
simple code or run code where I have already done the difficult
work for you. Let's see a few examples of how R works here before
moving on to today's material.

The text on a white background (like this!) is just plain text
that you can freely edit. Text that is on a grey background,
like the one below, is actual code. You can run it by clicking on
the green "play" button in the upper right-hand corner of the 
box. Everything here can be edited. Fill in your favorite number
below and then run the following code to create a new variable
with your number attached to it:

```{r}
fav_number <- 8
```

You should notice that your new variable shows up in the 
Environment tab in the upper right-hand corner of the RStudio
window. I will do my best to explain new R concepts when they
first arise. But, please ask questions in class about anything
you want to know more about!

## Taking the Mean

In the notes today we looked at taking the sample mean of a
dataset. Let's read in a dataset where we can take the mean of
one of the variable. Here, we have the code to read in a dataset
of the heights and birth-dates of players on rosters from the
2023/2024 Premier League season. A few heights are missing, and
we remove these before continuing:

```{r}
pl <- read_csv("../data/premier_league.csv")
pl <- filter(pl, !is.na(height))
pl
```

The variable `pl` contains the entire dataset. In order to 
access a specific column of the data, we use the dataset name
followed by a dollar sign and the name of the column of interest.
So, to get the heights, we would do `pl$height`. R has the 
function `mean()` that takes the sample mean of a set of numbers.
Below, we put these together to get the sample mean of the player
heights. The results are in centimeters.

```{r}
mean(pl$height)
```

How can we see this in terms of the random sampling that we have
in our theoretical notes? Well, that's actually a bit complex 
and philosophical. In a sense, we are not sampling from something
but rather observing the entire league. The easiest way to think
about this is to imagine that we are observing one set of players
that was generated be some complex genetic and social
interactions that led to these particular players being in the
Premier League during the 2023/2024 season. The number above
is an estimate of what the average height would be if we ran
back time and produced many different versions of our current
world. Yes, that's why I say that this gets quickly
philosophical! We will talk more about this throughout the 
semester.

We can take a mean of a specific part of the dataset by selecting
only some of the heights. For example, here is the code to
only select heights from the "AFC Bournemouth" team:

```{r}
mean(pl$height[pl$team == "AFC Bournemouth"])
```

Here are all of the teams in the dataset:

```{r}
table(pl$team)
```
In the code block below, find the average height of another 
team of your choosing.

```{r, question-01}
mean(pl$height[pl$team == "Chelsea FC"])
```

Think about how does this compare to the heights of the league
as a whole  and the AFC Bournemouth team specifically. Could 
you think of a way of characterizing the magnitude of these
differences?

## Simulation

Now, let's see how we can use R to simulate the relationships
between different random variables. We were able to derive some
of these theoretically. The simulations help to verify that our
calculations are correct and to help build further intuition.

The code below uses the `rnorm` function to generate a set of
25 samples from a normal distribution with mean 3 and standard
deviation 1.

```{r}
x <- rnorm(n = 25, mean = 3, sd = 1)
x
```

We can take the mean of these with the same `mean` function that
we used above. Try that in the code below:

```{r, question-02}
mean(x)
```

Compare your results with another student. You will see that 
your results are not exactly the same because you have different
random samples (in contrast, the means of the heights in the
previous section should have been exactly the same). 

We want to show that the distribution of the sample mean in this
case appears to follow a normal distribution with mean 3 and
variance 1 / n = 1/25. In order to do this we need to repeat the
previous simulation a large number of times and store the sample
mean each iteration. Here is the code to do that using a for
loop:

```{r}
N <- 100000
xbar <- rep(0, N)
for (j in seq_along(xbar))
{
  x <- rnorm(n = 25, mean = 3, sd = 1)
  xbar[j] <- mean(x)
}
```

Then, in order to see the simulated distribution, we can draw
a histogram with a curve on top of it. Note that R uses the
standard deviation to parameterize the normal distribution, so
we will use the square root of the expected variance: 1/5).

```{r}
hist(xbar, breaks = 50, probability = TRUE, col = "white")
curve(dnorm(x, 3, 1/5),
      from = min(xbar), to = max(xbar), add = TRUE, col = "red")
```

You should notice that the red line approximately follows the
empirical distribution given by the histogram. 

The central limit theorem says that we should have a similar
result for any sampling distribution with a finite mean and
finite variance. In the code below, run another simulation for
the sample mean of 25 i.i.d. samples. But this time, use a 
different distribution. Some options are: `runif`, `rbeta`, 
and `rgamma`. You may need to look up the parameter names for
the distribution.

```{r, question-03}
N <- 100000
xbar <- rep(0, N)
for (j in seq_along(xbar))
{
  x <- rbeta(n = 25, shape1 = 0.6, shape2 = 2)
  xbar[j] <- mean(x)
}
```

Now, create a histogram and overlay the expected normal
distribution. This may require looking up the properties of the
random variable.

```{r, question-04}
hist(xbar, breaks = 50, probability = TRUE, col = "white")
curve(dnorm(x, 0.6 / 2.6, 1/5 * sqrt((0.6 * 2) / (2.6^2 * (3.6)))),
      from = min(xbar), to = max(xbar), add = TRUE, col = "red")
```

If you have some extra time, see if reducing the sample size to
a smaller number begins to show the non-normality of the sample
mean. You should see that you can get some skew with smaller 
samples, but the CLT kicks in extremely fast.
