---
title: "Notebook 05"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

Let's read the keylogging data into R again. We will continue
to use this over the next weeks as it has a surprisingly large
number of different statistical applications.

```{r}
kl <- read_csv("../data/keylog.csv")
kl$key[is.na(kl$key)] <- " "
kl <- filter(kl, !is.na(gap1), !is.na(gap2))
kl
```

Today, we will look at two different samples, the amount of
time after typing a space bar until another key is typed, and
the amount of time after any non-space key and the next key 
is typed. We will pull these out as two variables called x and
y to match the notes; we also remove long gaps as per the 
discussion last time, and restrict ourselves to only looking
at the first copy task.

```{r}
x <- kl$gap2[kl$task == 1 & kl$key == " " & kl$gap2 < 1000]
y <- kl$gap2[kl$task == 1 & kl$key != " " & kl$gap2 < 1000]
```

We will also save the variable `n` and `m` to match those
from the notes:

```{r}
n <- length(x)
m <- length(y)
```

In the code below, compute the best guess point estimator for
the difference in the means of these two gaps:

```{r, question-01}
mean(x) - mean(y)
```

Now, let's build a two-sample confidence interval for the 
difference in means. To start, get the value t sub alpha,
which we call `tval` in R, for the appropriate degrees of
freedom derived on the worksheet for a 95% confidence interval:

```{r, question-02}
tval <- qt(1 - 0.05/2, df = nrow(kl))
tval
```

Next, compute the pooled variance in the code below, saving the
output as a variable called `sp`. 

```{r, question-03}
sp <- ((n - 1) * var(x) + (m - 1) * var(y)) / (n + m - 2)
sp
```

Finally, compute the confidence interval in the code below by 
applying the formula you constructed:

```{r, question-04}
c((mean(x) - mean(y)) - tval * sqrt(sp * (1/n + 1/m)),
  (mean(x) - mean(y)) + tval * sqrt(sp * (1/n + 1/m)))
```

Now, the easy part. Run the following code to apply the result
you have above with the built-in function in R: 

```{r}
t.test(x, y, conf.level = 0.95, var.equal = TRUE)
```

Note that the R function has an option that we set to indicate
that we want to assume that the variances are equal. Repeat
the procedure below but set `var.equal` to `FALSE`. This runs
a more complex algorithm that is (approximately) valid in the
case that the samples have different variances. Note how it
affects the results.

```{r, question-05}
t.test(x, y, conf.level = 0.95, var.equal = FALSE)
```

You should see that the confidence interval gets a bit larger.

Finally, let's apply the confidence interval for the ratio of
the variances. In the interest of time, I have typed out the
code below to generate the formula using the equation you should
have found on the worksheet:

```{r}
fval1 <- qf(1 - 0.05/2, df1 = n - 1, df2 = m - 1)
fval2 <- qf(0.05/2, df1 = n - 1, df2 = m - 1)

c(var(x) / var(y) * fval2, var(x) / var(y) * fval1)
```

Below, apply the function `var.test` to the two samples (no other
inputs required) in order to generate a 95%-confidence interval.
There are a number of different outputs; find the confidence
interval and verify that it matchs the code above.

```{r, question-06}
var.test(x, y)
```






