---
title: "Notebook 18"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

## 1. Bayesian Estimator for Binomial Probability

Following the class notes, here is a visualization of the 
posterior probability for the estimation of the proportion
p from a binomial sample. You can play around with the values
of the sample size (N) and the count of 1s (X), watching how
the posterior changes.

```{r}
N <- 5
X <- round(N * (0.25))
curve(dbeta(x, shape1 = X + 1, shape2 = N - X + 1),
      from = 0, to = 1, cex = 2, n = 1000)
abline(v = X / N, col = "red")
```

You should see that as N grows, the probability distribution
begins to move from a uniform distribution (the prior) to a
distribution concentrated on the mean.

## 2. Credible Region

Let's create an actual sample of data from a hypothetical
binomial distribution, where 25% of data are equal to 1 and
75% of the data are equal to 0.

```{r}
N <- 100
x <- rep(c(0, 1), c(N * (1-0.25), N * (0.25)))
x
```

The T-test function, as we know, can give a confidence interval
based on the T pivot statistic:

```{r}
t.test(x)
```

We can, alternatively, use our Bayesian approach to find a 
credible region for the value p:

```{r}
X <- sum(x)
cint <- c(
  qbeta(0.05/2, shape1 = X + 1, shape2 = N - X + 1),
  qbeta(1 - 0.05/2, shape1 = X + 1, shape2 = N - X + 1)
)
sprintf("95%% Confidence Interval: p ∈ [%.03f, %.03f]", cint[1], cint[2])
```

Notice that the value is very similar to the one based on the
T-test. The biggest advantage is that we have a much less 
complex way of describing the interpretation of the credible
interval. 
