---
title: "Notebook 14"
output:
  html_document:
    theme: simplex
    highlight: pygments
    css: "../css/note-style.css"
---

```{r, include=FALSE}
source("../funs/funs.R")
```

Let's briefly look at a well-known, built-in dataset in R that
records the fuel efficiency of a number of different types of 
cars. The data has the average miles per gallon driving in both
the city (lots of stops and slower speeds) and on the highway.

```{r}
mpg <- select(ggplot2::mpg, manufacturer, model, year, city = cty, highway = hwy)
mpg
```

What is the relationship between the city and highway fuel
efficiency? If we plot it, we see that the pattern closely 
follows a roughly linear relationship between the two:

```{r}
mpg |>
  ggplot(aes(city, highway)) +
    geom_point()
```

We can visually add a linear regression line with the following:

```{r}
mpg |>
  ggplot(aes(city, highway)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = 'y ~ x')
```

If we want to learn the actual parameters of the regression model,
the slope and intercept, we need to run the `lm` function 
directly and summarize the output. Here is what that looks like:

```{r}
model <- lm(highway ~ city, data = mpg)
summary(model)
```

The table gives the point estimate, the standard error (that's
the part of the confidence interval without the t-critical value),
the T-statistic testing the null hypothesis that the value is
zero, and the corresponding p-value. To directly get confidence
intervals, we can use the `confint` function as follows:

```{r}
confint(model, level = 0.02)
```

If you want to run a regression without an intercept, as on the
worksheet, we can do this by adding a negative one to the 
argument of the `lm` function:

```{r}
model <- lm(highway ~ city - 1, data = mpg)
summary(model)
```

This is not as common as running one with an intercept. We will
see more ways of extending this technique next time, as well as
seeing how to interpret the bottom of the summary table.




