%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUESTION

Let $X_1, \ldots, X_n \sim N(\mu_X, \sigma_X^2)$ be a random sample. Using
the results from the handout, construct a pivot statistic $T$ as a function
of $\bar{X}$, $S_X^2$, $\mu_X$, and $\sigma_X^2$ that has a distribution of
$t(n-1)$. Do not simplify.

% SOLUTION

We have the following, by simply plugging in the results from the handout:
\begin{align*}
T &= \frac{\frac{\mu_X - \bar{X}}{\sqrt{\sigma_X^2 / n}}}{\sqrt{\frac{(n-1)S^2_X}{\sigma_X^2 \cdot (n-1)}}}
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUESTION

Simplify the form of the $T$ statistic. It should no longer have any
$\sigma_X^2$ terms (in fact this is the whole point of this specific form). 
Try to write the solution with $(\mu - \bar{X})$ in the numerator and
everything else in the denominator.

% SOLUTION

Simplifying, we see that:
\begin{align*}
T &= \frac{\frac{\mu_X - \bar{X}}{\sqrt{\sigma_X^2 / n}}}{\sqrt{\frac{(n-1)S^2_X}{\sigma_X^2 \cdot (n-1)}}} \\
&= \frac{\mu_X - \bar{X}}{\sqrt{S^2_X / n}}
\end{align*}
As desired.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUESTION

Let $t_\alpha(k)$ be the tail probability of a T-distribution with $k$
degrees of freedom, just as we had with $z_\alpha$ on the handout. Following
the same procedure with the example on the handout, construct a confidence
interval with confidence level $(1 - \alpha)$ for $\mu_X$. Write the solution
as $\bar{X} \pm \Delta$ for some $\Delta$.

% SOLUTION

Starting with the pivot statistic, we have:
\begin{align*}
\Prob\left[ -t_\alpha(n-1) \leq T \leq t_\alpha(n-1) \right] &= 1 - \alpha \\
\Prob\left[ -t_\alpha(n-1) \leq \frac{\mu_X - \bar{X}}{\sqrt{S^2_X / n}} \leq t_\alpha(n-1) \right] &= 1 - \alpha \\
\Prob\left[ -t_\alpha(n-1) \cdot \sqrt{S^2_X / n} \leq (\mu_X - \bar{X}) \leq t_\alpha(n-1) \cdot \sqrt{S^2_X / n} \right] &= 1 - \alpha \\
\Prob\left[ \bar{X} - t_\alpha(n-1) \cdot \sqrt{S^2_X / n} \leq \mu_X \leq \bar{X} + t_\alpha(n-1) \cdot \sqrt{S^2_X / n} \right] &= 1 - \alpha \\
\end{align*}
Which we can write as the following:
\begin{align*}
\bar{X} &\pm t_\alpha(n-1) \cdot \sqrt{\frac{S^2_X}{n}}
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUESTION

We will go back to the story about the fish. Say we sample $25$ fish and
have a sample mean of $12.1$ centimeters and a sample variance of $6$
centimeters squared. Given that $t_{0.01/2}(24)$ is approximately equal
to $2.797$, derive the confidence interval for the mean.

% qt(1 - 0.01/2, df = 24)
% 2.797 * sqrt(6/24)

% SOLUTION

We have:
\begin{align*}
12.1 &\pm \left[ 2.797 \cdot \sqrt{\frac{6}{24}} \right] \\
12.1 &\pm 1.3985
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUESTION

Let $C_k \sim \chi^2(k)$ for every integer $k$. Use Chebychev's Inequality to show
that for any $\epsilon > 0$, we have:
\begin{align*}
\lim_{k \rightarrow \infty} \mathbb{P} \left[ |C_k/k - 1| \geq \epsilon \right] = 0
\end{align*}
In this case we say that $C_k$ limits in probability to 1, written as
$C_k \rightarrow_P 1$.

% SOLUTION

TODO

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUESTION

Let $Y_n \rightarrow_P y$ for a constant $y$, $f$ is a real-valued function that
is invertable around the neighborhood of $y$, and $X$ is another random variable.
Then, Slutsky's Theorem says that $g(Y_n) \cdot X$ limits in probability to
$g(y) \cdot X$. Use this to show that the $T$ distribution limits to the standard
normal as the degrees of freedom limit to infinity.

% SOLUTION

TODO
